---
title: "Reading In MRIcloud Data"
author: "Brian Caffo"
date: "April 18, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading in T1 Volumetric output and working with it

- MRICloud formats structrual and functional output in
a text format
- In this tutorial, we discuss how to text process MRICloud
output
- We're creating a set of utilies for working with MRICloud output
- We'll discuss T1 output today

## Example T1 output

```{r}
##working in the directory where the ROIs are at
roiDir = "../data/"
fileList = dir(roiDir)
## grab the subject id
subjects = sapply(strsplit(fileList, "_"), function(x) x[1])
```

## These have to be downloaded from github

```{r, eval=FALSE}
## Only if you don't have the package installed
#source("https://neuroconductor.org/neurocLite.R")
#neuro_install("MRIcloudT1volumetrics")
require(devtools)
devtools::install_github("bcaffo/MRIcloudT1volumetrics")
```

## Read in a subject 

- This just grabs volume
- Uses the readSubject function from `MRIcloudT1volumetrics`

```{r, message = FALSE}
library(magrittr); library(dplyr); library(MRIcloudT1volumetrics) 
## Change this to where the data directory is
roiDir = "~/sandboxes/MRIcloudTutorial/data/"
fileList = dir(roiDir)
fullPath = paste(roiDir, fileList[1], sep = "")
dat = readSubject(fullPath)
length(dat)
```

10 for two types and 5 levels, each type/level combo is a list element

## Data is a list

```{r}
dat[[1]]
```

## Turns the whole subject into a data frame
```{r}
dat = subject2df(dat)
##makes printing easier gets rid of the id variable which is the filename
dat$rawid = NULL 
head(dat, 5)
```

## Adds ICV variable
```{r, warning = FALSE, message = FALSE}
dat = addSubjectICV(dat)
head(dat, 5)
```

## Adds Subject TBV variable
```{r, warning = FALSE, message = FALSE}
dat = addSubjectTBV(dat)
head(dat, 5)
```

## Doing things manually
```{r}
fullData = readLines(fullPath)
##print out part of it
cat(substr(fullData[11 : 18], start = 20, stop = 60), sep = "\n")
```

## Parsing the text data
```{r}
toparse = paste(fullData[11 : 18], "", collapse = "\n")
level1 = read.table(textConnection(toparse),
           fill = TRUE,
           stringsAsFactors = FALSE)
head(level1[,-1], 3)
```

## Renaming column names
```{r}
# could get the column names here
scan(textConnection(fullData[2]), what = "character")
names(level1) = c("id", "obj", "vol", "min", "max", "mean", "std")
level1$id = NULL
```

## Getting ICV and TBV
```{r}
library(dplyr); library(magrittr)
level1 %>% mutate(icv = sum(vol * (obj != "CSF")), 
                  tbv = sum(vol)) %>%
  select(-min, -max, -mean, -std)
```

## Other things to try
- Searching the file for headings to parse (this is what readSubject does)
- Plotting the data (another lecture)
- Switching between long and short formats (another lecture)
- Modeling the data (another lecture)