---
title: "ggplot"
author: "Brian Caffo"
date: "8/22/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get ggplot
```{r, eval=FALSE}
install.packages("ggplot2")
library(ggplot2)
```

## Get MRIcloudT1volumetrics
```{r, eval=FALSE}
## Only if you don't have the package installed
#source("https://neuroconductor.org/neurocLite.R")
#neuro_install("MRIcloudT1volumetrics")
require(devtools)
devtools::install_github("bcaffo/MRIcloudT1volumetrics")
```


## Load up the libraries
```{r, eval = TRUE, message = FALSE}
# do this first if needed
#devtools::install_github("bcaffo/kirby21.mricloud")
library(magrittr); library(dplyr); 
library(MRIcloudT1volumetrics); library(kirby21.mricloud)
library(tidyr); library(ggplot2)
```

## Organize the list of ids (manually)
```{r}
#get the list of filenames
fileNames = get_txt_filenames(visits = 1, 
                              file_types =  "volumes_corrected")
```

## Read in the data
```{r}
#read them in
dat = readFileList(fileNames$filename, 
                   level= 1, 
                   type = 1, 
                   idList = fileNames$Subject_ID) %>%
  select(id, roi, volume)
head(dat, 10)
```

## Average volume with each region

```{r}
smry = dat %>% group_by(roi) %>% 
  summarise(roiMean = mean(volume),
            roiSD = sd(volume), 
            n = n()
            ) %>% mutate(roi = as.factor(roi))
smry
```

## Using ggplot

```{r, warning = FALSE, message = FALSE, echo = TRUE}
g = ggplot(smry, aes(x = roi, y = roiMean))
g = g + geom_col()
```

## 
```{r, echo = FALSE}
g
```

## Changing the axis angle
`theme` controls the non data aspects of the plot

```{r}
g = g + theme(axis.text.x=element_text(angle=90,hjust=1)) 
```

##

```{r, echo = FALSE}
g
```


## Creating a stacked bar chart example

Parameters in the `aes` function control aesthetics 

```{r}
g = ggplot(dat, aes(x = as.factor(id), 
                    y = volume, 
                    fill = as.factor(roi))
           )
g = g + geom_col()
```

##

```{r, echo = FALSE}
g
```

## Merge with the demographic data
```{r}
dat = right_join(kirby21.base::kirby21_demog, 
                 dat, by = c("Subject_ID" = "id"))
head(dat, 10)
```

## Line plot by age
```{r}
g = ggplot(dat, aes(x = Age, y = log10(volume), col = roi))
g = g + geom_line()
```

##

```{r, echo = FALSE}
g
```

## Gender by roi
```{r}
smryGender = group_by(dat, Sex, roi) %>% 
  summarise(roiMean = mean(volume))  %>% 
  drop_na()
g = ggplot(smryGender, aes(x = Sex, y = roiMean, fill = roi)) + geom_col()
```

##

```{r, echo = FALSE}
g
```