---
title: "dplyr for ROI data wrangling"
author: "Brian Caffo"
date: "April 14, 2017"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## dplyr

- `dplyr` is a package for manipulating data in R
- It is an essential tool for working with ROI structural data
- Requires that you get your data into a *data frame* before starting
- 

## magrittr

- `magrittr` is a package that is very useful along with dplyr
- "The magrittr package offers a set of operators which promote semantics that will improve your code" 
- ` install.packages("magrittr")` 
- Here's what it does. It converts code like this:


```{r, eval = FALSE, echo = TRUE}
a = f(b)
a = g(a, c)
a = h(a, d)
```

with code like this

```{r, eval = FALSE, echo = TRUE}
a = f(b) %>% g(c) %>% h(d)
```

## dplyr
- Back to `dplyr`. This is a package for manipulating and wrangling data frames
- `dplyr` includes a collection of verbs for manipulating data, we'll go through them one at a time
- [https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html](https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html)
- The main ones for me are: `filter`, `select`, `mutate`, `summarise`


## Example data

```{r,echo=FALSE}
source("~/sandboxes/MRIcloudT1volumetrics/R/readSubject.R")
source("~/sandboxes/MRIcloudT1volumetrics/R/splitFileName.R")
source("~/sandboxes/MRIcloudT1volumetrics/R/subject2df.R")
source("~/sandboxes/MRIcloudT1volumetrics/R/utils.R")
```

```{r}
library(dplyr)
roiDir = "data/"
fileList = dir(roiDir)
##includes all types and levels
dat = readSubject(paste(roiDir, fileList[1], sep = "")) %>% subject2df
dim(dat)
## just look at the first level
filter(dat, type == 1, level == 1)
```

## Create ICV normalized volumes
```{r}
dat %>%  mutate(norm.v
```

