---
title: "dplyr for ROI data wrangling"
author: "Brian Caffo"
date: "April 14, 2017"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

## dplyr

- `dplyr` is a package for manipulating data in R
- It is an essential tool for working with ROI structural data
- Requires that you get your data into a *data frame* before starting

## magrittr

- `magrittr` is a package that is very useful along with dplyr
- "The magrittr package offers a set of operators which promote semantics that will improve your code" 
- ` install.packages("magrittr")` 
- Here's what it does. It converts code like this:


```{r, eval = FALSE, echo = TRUE}
a = f(b)
a = g(a, c)
a = h(a, d)
```

with code like this

```{r, eval = FALSE, echo = TRUE}
a = f(b) %>% g(c) %>% h(d)
```

## dplyr
- Back to `dplyr`. This is a package for manipulating and wrangling data frames
- `dplyr` includes a collection of verbs for manipulating data, we'll go through them one at a time
- [Here's a link to the best docs](https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html)
- The main ones for me are: `filter`, `select`, `mutate`, `summarise`


## Example data

- We'll work with some example data in the github repo under `data/`
- Contains two subjects' from the Kirby21 dataset

```{r}
dir("../data")
```


## Utility functions 

```{r,echo=TRUE}
source("~/sandboxes/MRIcloudT1volumetrics/R/readSubject.R")
source("~/sandboxes/MRIcloudT1volumetrics/R/splitFileName.R")
source("~/sandboxes/MRIcloudT1volumetrics/R/subject2df.R")
source("~/sandboxes/MRIcloudT1volumetrics/R/utils.R")
```

## Load in dplyr and read in the data
```{r, warning = FALSE, message = FALSE, echo = TRUE}
library(dplyr)
roiDir = "../data/"
fileList = dir(roiDir)
dat = readSubject(paste(roiDir, fileList[1], sep = "")) %>% 
    subject2df()
```

## The data set at this point

```{r}
head(dat, 10)
```

## Get rid of the id column

- Minus removes columns
- Otherwise `select` will keep the columns that you specify

```{r}
dat = select(dat, -rawid)
head(dat, 10)
```

## Using filter

- `filter` selects rows of a dataframe based on criteria

```{r, echo= TRUE}
dat  = filter(dat, type == 1, level == 1)
dat
```

## Create ICV normalized volumes

- There's already a function to do this in the `MRIcloudt1Volumetrics` repo
-However, let's do it manually

```{r}
dat = mutate(dat, icv = sum(volume))
dat
```

## Why magrittr is so useful

Everything we've done so far

```{r, eval = FALSE, echo = TRUE}
dat = readSubject(paste(roiDir, fileList[1], sep = "")) %>% 
    subject2df() %>%
    filter(type == 1, level == 1) %>%
    mutate(icv = sum(volume))
```

## Let's return to the full data
```{r}
fullData = readSubject(paste(roiDir, fileList[1], sep = "")) %>% 
        subject2df()
head(fullData)
```


## Summarise

- Question, what's the volumetric sum at each level and type?
- `summarise` creates summaries
- `group_by` operates functions over groups

## 

```{r}
fullData %>% group_by(type, level) %>% 
    summarise(nvol = n(), sum = sum(volume), mean = mean(volume)) %>%
    as.data.frame() #just for display
```





