---
title: "Long versus wide format"
author: "Brian Caffo"
date: "April 19, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

## Load in the sample data

```{r,echo=FALSE}
source("~/sandboxes/MRIcloudT1volumetrics/R/readSubject.R")
source("~/sandboxes/MRIcloudT1volumetrics/R/splitFileName.R")
source("~/sandboxes/MRIcloudT1volumetrics/R/subject2df.R")
source("~/sandboxes/MRIcloudT1volumetrics/R/utils.R")
```

```{r, eval=FALSE}
source("readSubject.R")
source("splitFileName.R")
source("subject2df.R")
source("utils.R")
```

## reading in the data

```{r, warning = FALSE, message = FALSE}
library(dplyr)
roiDir = "../data/"
fileList = dir(roiDir)
dat = NULL
for (i in 1 : 2){
    tempDat = readSubject(paste(roiDir, fileList[i], sep = "")) %>% 
        subject2df() %>%
        filter(level == 1, type == 1) %>%
        mutate(subject = i) %>% 
        select(roi, volume, subject)
    dat = rbind(dat, tempDat)
}
```

## Long versus wide format

- An important organizational principle for ROI data is long versus wide format
- Long formats are often useful for analysis software, like multilevel models
- Wide formats are more compact and useful for data manipulation


## Long format
```{r, echo = FALSE}
library(tidyr)
dat
```


## Wide format

Showing the first 5 columns (otherwise I run out of room)

```{r}
spread(dat, roi, volume)[,1:5]
```

## Conversions

- R has several wide to long and vice-versa converters
- Notable is the `reshape2` pacage and `tidyr`
- Also is the `reshape` function (which has nothing to do with `reshape2`)
- We're recommending `tidyr` 
    - `gather` converts from wide to long
    - `spread` converts from long to wide

## long to wide again

```{r}
wide = spread(dat, roi, volume)
```

- `dat` the data frame
- `roi` the name of column variables
- `volume` the value of the columns

## Note
- Note you can't have duplicates since `spread` then can't figure out what to do
- The following (not evaluated) gives and error since there's duplicate subjects

```{r, echo = TRUE, eval = FALSE}
tempDat = rbind(dat, dat)
spread(tempDat, roi, volume
```


## wide back to long

```{r}
gather(wide, roi, volume, CSF:Telencephalon_R) %>%
    arrange(subject)
```


